Fully Automatic Loop-free Program Synthesis by Bounded Model Checking

### 2014-01-20 paper/paper.tex
Added references to a couple of figures

Additionally, \newC programs may use nondeterministic values, assumptions
and arbitrary-width types.

@@

Our encoding is the following:
%
\begin{itemize}
 \item $I$ is the set of $N$-tuples of 32-bit bitvectors.  This is written in \newC as the type \verb|int[N]|.
 \item $O$ is the set of $M$-tuples of 32-bit bitvectors, which is written in \newC as the type \verb|int[M]|.
 \item $\sigma$ is a pure function with type $I \times O \rightarrow \mathrm{Bool}$.  The \newC signature of this function is
 \verb|int check(int in[N], int out[M])|. This function is the only component supplied
 by the user.
 \item $P$ is written in a simple RISC-like language $\mathcal{L}$.  Programs in $\mathcal{L}$
 have type $I \rightarrow O$ and
 are represented in \newC as objects of type \verb|prog_t|.
 \item We supply an interpreter for $\mathcal{L}$ which is written in \newC.  The type of
 this interpreter is $(I \rightarrow O) \times I \rightarrow O$ and the \newC signature is \\
 \verb|void exec(prog_t p, int in[N], int out[M])|.  Here, \verb|out| is an output parameter.
\end{itemize}

@@

\begin{figure}
\begin{tikzpicture}[scale=0.9, transform shape, node distance=2cm, auto]
\tikzstyle{decision} = [diamond, draw, fill=blue!20, 
    text width=4.5em, text badly centered, node distance=3cm, inner sep=0pt]
\tikzstyle{block} = [rectangle, draw, fill=blue!20, 
    text width=5em, text centered, rounded corners, minimum height=4em]
\tikzstyle{line} = [draw, -latex']
\tikzstyle{cloud} = [draw, ellipse,fill=red!20, node distance=3cm,
    minimum height=2em]

@@

\begin{figure}
\begin{center}
\begin{minipage}[t]{0.45\linewidth}
\begin{algorithm}[H]
\caption{\sc 2Sum
 \label{alg:2sum}}
\begin{algorithmic}
\Let{$s$}{$a+b$}
\Let{$b'$}{$s - a$}
\Let{$a'$}{$s - b'$}
\Let{$\delta b$}{$b - b'$}
\Let{$\delta a$}{$a - a'$}
\Let{$t$}{$\delta a + \delta b$}
\Ensure{$a + b$ = $s + t$ exactly}
\end{algorithmic}
\end{algorithm}
\end{minipage}
\hfill
\begin{minipage}[t]{0.45\linewidth}
\begin{algorithm}[H]
\caption{\sc Fast2Sum
 \label{alg:fast2sum}}
\begin{algorithmic}
\Require{$a \ge b$}
\Let{$s$}{$a + b$}
\Let{$z$}{$s - a$}
\Let{$t$}{$b - z$}
\Ensure{$a + b$ = $s + t$ exactly}
\end{algorithmic}
\end{algorithm}
\end{minipage}
\end{center}

@@

To demonstrate the capabilities of \newC as an implementation language, we synthesise a
non-trivial floating-point program.  For this we chose the classic {\sc 2Sum} algorithm for computing
exact sums of floating-point numbers.  The original {\sc 2Sum} algorithm is given as Alg.~\ref{alg:2sum}.
In the case that $a \ge b$, exact addition can be implemented in fewer instructions -- this is the
{\sc Fast2Sum} algorithm~\cite{fast2sum}.  We used {\sc 2sum} as our specification, and
added an assumption that $a \ge b$.  {\sc Kalashnikov} is able to synthesise and verify the
code for {\sc Fast2Sum} from this specification in 396.07\,s, which we believe to be the first time
this has been achieved.  {\sc 2Sum} was proved to be minimal in~\cite{fast2sum}, but
we believe our synthesis constitutes the first proof that {\sc Fast2Sum} is
minimal under the assumption that $a \ge b$.

>>>>

Additionally, \newC programs may use nondeterministic values, assumptions
and arbitrary-width types.  Two example \newC programs are shown 
in Fig.~\ref{fig:c-}.

@@

Our encoding is the following:
%
\begin{itemize}
 \item $I$ is the set of $N$-tuples of 32-bit bitvectors.  This is written in \newC as the type \verb|int[N]|.
 \item $O$ is the set of $M$-tuples of 32-bit bitvectors, which is written in \newC as the type \verb|int[M]|.
 \item $\sigma$ is a pure function with type $I \times O \rightarrow \mathrm{Bool}$.  The \newC signature of this function is
 \verb|int check(int in[N], int out[M])|. This function is the only component supplied
 by the user.
 \item $P$ is written in a simple RISC-like language $\mathcal{L}$, described in Fig.~\ref{fig:l-language}.  Programs in $\mathcal{L}$
 have type $I \rightarrow O$ and
 are represented in \newC as objects of type \verb|prog_t|.
 \item We supply an interpreter for $\mathcal{L}$ which is written in \newC.  The type of
 this interpreter is $(I \rightarrow O) \times I \rightarrow O$ and the \newC signature is \\
 \verb|void exec(prog_t p, int in[N], int out[M])|.  Here, \verb|out| is an output parameter.
\end{itemize}

@@

\begin{figure}[h]
\begin{tikzpicture}[scale=0.9, transform shape, node distance=2cm, auto]
\tikzstyle{decision} = [diamond, draw, fill=blue!20, 
    text width=4.5em, text badly centered, node distance=3cm, inner sep=0pt]
\tikzstyle{block} = [rectangle, draw, fill=blue!20, 
    text width=5em, text centered, rounded corners, minimum height=4em]
\tikzstyle{line} = [draw, -latex']
\tikzstyle{cloud} = [draw, ellipse,fill=red!20, node distance=3cm,
    minimum height=2em]

@@

\begin{figure}[h]
\begin{center}
\begin{minipage}[t]{0.45\linewidth}
\begin{algorithm}[H]
\caption{\sc 2Sum
 \label{alg:2sum}}
\begin{algorithmic}
\Let{$s$}{$a+b$}
\Let{$b'$}{$s - a$}
\Let{$a'$}{$s - b'$}
\Let{$\delta b$}{$b - b'$}
\Let{$\delta a$}{$a - a'$}
\Let{$t$}{$\delta a + \delta b$}
\Ensure{$a + b$ = $s + t$ exactly}
\end{algorithmic}
\end{algorithm}
\end{minipage}
\hfill
\begin{minipage}[t]{0.45\linewidth}
\begin{algorithm}[H]
\caption{\sc Fast2Sum
 \label{alg:fast2sum}}
\begin{algorithmic}
\Require{$a \ge b$}
\Let{$s$}{$a + b$}
\Let{$z$}{$s - a$}
\Let{$t$}{$b - z$}
\Ensure{$a + b$ = $s + t$ exactly}
\end{algorithmic}
\end{algorithm}
\end{minipage}
\end{center}

@@

To demonstrate the capabilities of \newC as an implementation language, we synthesise a
non-trivial floating-point program.  For this we chose the classic {\sc 2Sum} algorithm for computing
exact sums of floating-point numbers.  The original {\sc 2Sum} algorithm is given as Alg.~\ref{alg:2sum}.
In the case that $a \ge b$, exact addition can be implemented in fewer instructions -- this is the
{\sc Fast2Sum} algorithm~\cite{fast2sum}, shown in Alg.~\ref{alg:fast2sum}.  We used {\sc 2sum} as our specification, and
added an assumption that $a \ge b$.  {\sc Kalashnikov} is able to synthesise and verify the
code for {\sc Fast2Sum} from this specification in 396.07\,s, which we believe to be the first time
this has been achieved.  {\sc 2Sum} was proved to be minimal in~\cite{fast2sum}, but
we believe our synthesis constitutes the first proof that {\sc Fast2Sum} is
minimal under the assumption that $a \ge b$.
